### 2017-05-09 paper.tex
Merge remote-tracking branch 'origin/master'
# Conflicts:
#	paper.pdf
Move new content from paper.tex to paper.Rnw

\documentclass{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,float}
\usepackage{multirow,setspace}
\usepackage{enumerate}
\usepackage{booktabs}
\usepackage{placeins}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}

@@

\setlength{\marginparwidth}{2cm}
\usepackage{Sweave}
\begin{document}
\input{paper-concordance}
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
The data set contains a variety of measures from severe weather events in the United States from 1996-2016. 
For this project we focused on the deaths directly attributable to the event.
Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.

@@

\noindent Bayarri, Berger, and Datta (2008) suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior

@@

\begin{thebibliography}{1}
  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
    Accessed April 2017.
   \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
\end{thebibliography}
\clearpage
\appendix
\section{Code}
\label{a:code}
This appendix includes the code used to implement the models.

\section{Derivations}
\label{a:derivation}
This appendix will include details on the calculations required to derive our models.
\subsection{Zero Inflated Negative Binomial Derivation}
\label{a:dZINB}
The likelihood for the ZINB is 
\begin{align*}
\mathcal{L}(X|\sigma, p, r) &= \prod_{i=1}^N \sigma I_{X=0}(X_i) + (1-\sigma)\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\\
\intertext{For ease of notation let $Z$ be the number of zero values in $X$, and $N$ be the total number of observations.}
&=\prod_{X_i=0}\left(\sigma + (1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
&=\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^n\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)
\end{align*}
As mentioned in section \ref{ss:mZiNBinom} we take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. Therefore my joint posterior is:
\begin{align*}
p(r,\sigma, p|X)\propto&\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
\intertext{I am now going to take the log of the posterior because it helps with computation}
\ln\left(p(r,\sigma, p|X)\right)\propto&Z\ln\left(1 + (1/\sigma-1)(1-p)^r\right) + Z\ln(\sigma) + (N-Z)\ln(1-\sigma)+(N-Z)r\ln(1-p)\\
&+\sum_{i=1}^NX_i\ln(p)-\ln(r)/2-N\ln(\Gamma(r))+\sum_{i=1}^N\ln(\Gamma(r+X_i))
%\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
\end{align*}

>>>>

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,float}
\usepackage{multirow,setspace}
\usepackage{enumerate}
\usepackage{booktabs}
\usepackage{placeins}
\usepackage{cite}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}

@@

\setlength{\marginparwidth}{2cm}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
For this project we focused on the deaths directly attributable to the event.
Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.

@@

\noindent Bayarri, Berger, and Datta (2008) \cite{ZIP} suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior

@@

\begin{thebibliography}{2}
  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
    Accessed April 2017.
  \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
\end{thebibliography}
\clearpage
\appendix
\section{Code}
\label{a:code}
This appendix includes the code used to implement the models.

\section{Derivations}
\label{a:derivation}
This appendix will include details on the calculations required to derive our models.
\subsection{Zero Inflated Negative Binomial Derivation}
\label{a:dZINB}
The likelihood for the ZINB is 
\begin{align*}
\mathcal{L}(X|\sigma, p, r) &= \prod_{i=1}^N \sigma I_{X=0}(X_i) + (1-\sigma)\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\\
\intertext{For ease of notation let $Z$ be the number of zero values in $X$, and $N$ be the total number of observations.}
&=\prod_{X_i=0}\left(\sigma + (1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
&=\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^n\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)
\end{align*}
As mentioned in section \ref{ss:mZiNBinom} we take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. Therefore my joint posterior is:
\begin{align*}
p(r,\sigma, p|X)\propto&\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
\intertext{I am now going to take the log of the posterior because it helps with computation}
\ln\left(p(r,\sigma, p|X)\right)\propto&Z\ln\left(1 + (1/\sigma-1)(1-p)^r\right) + Z\ln(\sigma) + (N-Z)\ln(1-\sigma)+(N-Z)r\ln(1-p)\\
&+\sum_{i=1}^NX_i\ln(p)-\ln(r)/2-N\ln(\Gamma(r))+\sum_{i=1}^N\ln(\Gamma(r+X_i))
\end{align*}
